---
# TITLE & AUTHOR
title: "Diss"
subtitle: "Current Status"
author: "Joschka Schwarz"
institute: "Hamburg University of Technology"
date: today
date-format: "dddd, D[<sup style='font-size:65%;font-style:italic;'>th</sup>] [of] MMMM YYYY"
---

## Identity matching {data-name="DATA"}

Linking Developer with startup data

::: {.absolute top=30% left=10% width=80%}
![](img/data_sources_v2.svg)
:::

## Research Question {data-name="Research Question"}

When you click the **Render** button a document will be generated that includes:

-   Content authored with markdown
-   Output from executable code

## Code {data-name="Forschung und Transfer"}

When you click the **Render** button a presentation will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```

## Why {data-name="Dekanate"}

When you click the **Render** button a presentation will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```

## Same {data-name="International"}

When you click the **Render** button a presentation will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```

## Different {data-state="hide-menubar"}

When you click the **Render** button a presentation will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```
